Augmenting the DonorsChoose.org Corpus for Meta-Learning \thanks{This publication emanated from research conducted with the financial support of Science Foundation Ireland (SFI) under Grant Number 13/RC/2106.}

### 2019-03-31 doc/paper/main.tex
paper/main.tex: Grammar and wording details

\begin{abstract}
The DonorsChoose.org dataset of past donations provides a big and feature-rich corpus of users and items. The dataset matches donors to projects in which they might be interested in and hence is intrinsically about recommendations. Due to the availability of detailed item-, user- and transaction-features, this corpus represents a suitable candidate for meta-learning approaches to be tested. This study aims at providing an augmented corpus for further recommender systems studies to test and evaluate meta-learning approaches. In the augmentation, metadata of collaborative and content-based filtering techniques is amended to the corpus. It is further extended with aggregated statistics for users and transactions and an exemplary meta-learning experiment. The performance in the learning subsystem is measured via the recall of recommended items in a Top-N test set. The augmented dataset and the source code are released into the public domain at \href{https://github.com/BeelGroup/Augmented-DonorsChoose.org-Dataset}{GitHub:BeelGroup/Augmented-DonorsChoose.org-Dataset}.

@@

One challenge for researchers interested in meta-learning is that suitable datasets are time-consuming and cumbersome to construct. Creating a system of individual machine learning experiments and amending their metadata to the original corpus takes time and resources away from working with the actual meta-learners. Multiple examples of manual and repetitious data augmentation may be found in the scientific literature, for instance in~\cite{CUNHA2018128,DBLP:journals/corr/abs-1805-12118,Ekstrand:2012:RFP:2365952.2366002}. Furthermore, the choice of the metric for computing metadata in the system of single machine learners is important. Using the RMSE, as is done in some of the aforementioned publications, disallows using content-based methods.

Our objective is to address the data augmentation part in a comprehensive and reproducible~\cite{UMUAI:TowardsReproducibilityInRecSysBeel2016} way as to allow researchers to focus on the actual meta-learning task. We are presenting a public domain dataset which to our best knowledge has not been used for meta-learning before in the scientific community, yet due to its feature-rich and vast corpus is a suitable candidate. The raw dataset is augmented with metadata of several machine learning algorithms from a variety of fields (cf.~\cite{DBLP:journals/corr/abs-1805-12118}). In place of a metric, the recall in a Top-N set is used as it can be computed for a wide range of algorithms. The dataset is further amended with aggregated statistics about users and items and metadata of an exemplary experiment.

@@

The learning subsystem comprises two collaborative filtering techniques and two content-based filtering techniques. The performance measures of the single algorithms are amended to the table. Next, aggregated user and item statistics are calculated and added to the table. Lastly, four exemplary meta-learners are applied and the performance measures of the individual approaches are added to the corpus.

@@

\subsection{Dataset Preparation}
To prepare the dataset for evaluations, duplicate interactions are merged as the recommender systems shall not recommend items to which the user has already donated to. Internal stop words in description strings of items are stripped. Furthermore, transactions containing no information about the user's location are dropped. Thereby enabling the consistent usage of a user's location information in new approaches.

Additionally, users having interacted less than twice are removed as well. Although, this limits the target group, the step is unavoidable as the employed validation process requires at least one interaction per user for testing and one for training. The evaluation requires user profiles and hence training-transactions in order to recommend new items which in turn need to be validated against test-transactions. Lastly, the donated amount is transformed to a transaction strength in the range of unity to five in analogy to a rating score. This step reduces the influence of outliers respectively unusual donated amounts.

\subsection{Learning Subsystem}
The learning subsystem is a recommender system itself. SVD and KNN are used as representatives for collaborative filtering techniques and TF-IDF and word-embedding for content-based filters~\cite{scikit-learn,rehurek_lrec}. Hereby, the otherwise intrinsic inductive bias is reduced. The word-embedding relies on fastText and utilizes pre-trained vectors~\cite{DBLP:journals/corr/BojanowskiGJM16}. Vector representations of items are the normalized sum of the vectors of the normalized embeddings of the individual words. In order to work with the data in an efficient way, transactions are sampled. For this study a sample-size of $100,000$ was chosen. This represents a fair balance between required computing time and ability to generalize.

@@

A classical 5-fold cross-validation on the users is applied for training the collaborative filters. An ordering is assumed to be implied by the magnitudes of the values in the reconstructed matrix of the decomposition respectively the distance to the neighbor. The validation process of the content-based approaches uses leave-one-out. The ranking is performed using the cosine similarity of an item's vector and the user's profile.

@@

\subsection{Statistics and Application of Learning Subsystem}
The meta-learning system processes the metadata from the learning subsystem and further statistics. Its performance is evaluated via a holdout split on the transaction table using $80\%$ for training and the rest for testing. Transactions are expanded to contain more detailed information about the time, the user's location and aggregated statistics about the mean of the categories, location and interaction-strength of items with which the user has interacted with. Statistics are solely based on the training set. Values for unknown users are filled with the mean in the training data.

Four different meta-learning approaches are discussed which address the Algorithm Selection Problem (ASP) via a switching hybrid ensemble. The aim of the system is to predict the algorithm with the lowest recall-position for each transaction respectively each row in~\autoref{tab:IllustrativeTable}. The overall performance of a meta-learner is measured using the mean recall-position for when the recommended algorithm for each transaction is used.

@@

\begin{figure}[ht]
	\centering
	\subcaptionbox{Interactions Frequency\label{fig:InteractionFrequency}}{
		\includegraphics[width=0.45\textwidth,height=\textheight,keepaspectratio]{{{../res/number_interactions}}}
	}
	\hspace{2em}
	\subcaptionbox{Interaction Strength\label{fig:InteractionStrength}}{
		\includegraphics[width=0.45\textwidth,height=\textheight,keepaspectratio]{{{../res/distribtion_ratings}}}
	}
	\caption{The frequency of the number of interactions per user without outliers is shown in~\subref{fig:InteractionFrequency} and the distribution of the transformed interaction strength is shown in~\subref{fig:InteractionStrength}.}\label{fig:Interactions}
\end{figure}

@@

That the stacking ensemble performs best is unsurprising as it is provided significantly more information than the other algorithms. The performance of the gradient booster is encouraging as it successfully addresses the ASP. From the poor performance of the user-clustering it can be reasoned that similar transactions do not necessarily yield similar results using the same recommender algorithm. Hence, the similarity of meta-feature does not imply that the performance of recommendations is similar.

@@

\section{Conclusion}
The study provides an extensively augmented dataset based on the transaction data published by DonorsChoose.org. The corpus is amended with metadata of individual recommender algorithms. Collaborative and content-based filtering techniques are used in the process and their performance is evaluated via the recall in a Top-N test set. Aggregated user and item statistics are amended to the table. Metadata of four switching hybrid ensemble meta-learners is amended to the dataset. The augmented public domain dataset lays the groundwork that future evaluations of existing and novel meta-learning approaches can build upon.

>>>>

\begin{abstract}
The DonorsChoose.org dataset of past donations provides a big and feature-rich corpus of users and items. The dataset matches donors to projects in which they might be interested in and hence is intrinsically about recommendations. Due to the availability of detailed item-, user- and transaction-features, this corpus represents a suitable candidate for meta-learning approaches to be tested. This study aims at providing an augmented corpus for further recommender systems studies to test and evaluate meta-learning approaches. In the augmentation, metadata of collaborative and content-based filtering techniques is amended to the corpus. It is further extended with aggregated statistics of users and transactions and an exemplary meta-learning experiment. The performance in the learning subsystem is measured via the recall of recommended items in a Top-N test set. The augmented dataset and the source code are released into the public domain at \href{https://github.com/BeelGroup/Augmented-DonorsChoose.org-Dataset}{GitHub:BeelGroup/Augmented-DonorsChoose.org-Dataset}.

@@

One challenge for researchers interested in meta-learning is that suitable datasets are time-consuming and cumbersome to construct. Creating a system of individual machine learning experiments and amending their metadata to the original corpus takes time and resources away from working with the actual meta-learners. Multiple examples of manual and repetitious data augmentation may be found in the scientific literature, for instance in~\cite{CUNHA2018128,DBLP:journals/corr/abs-1805-12118,Ekstrand:2012:RFP:2365952.2366002}. Furthermore, the choice of the metric for computing metadata of single machine learners is important. Using the RMSE, as is done in some of the aforementioned publications, disallows using content-based methods.

Our objective is to address the data augmentation part in a comprehensive and reproducible~\cite{UMUAI:TowardsReproducibilityInRecSysBeel2016} way as to allow researchers to focus on the actual meta-learning task. We are presenting a public domain dataset which to our best knowledge has not been used for meta-learning before in the scientific community, yet due to its feature-rich and vast corpus is a suitable candidate. The raw dataset is augmented with metadata of several machine learning algorithms from a variety of fields (cf.~\cite{DBLP:journals/corr/abs-1805-12118}). In place of a metric, the recall in a Top-N set is used as it can be computed for a wide range of algorithms. The dataset is further amended with aggregated statistics about users and items, and metadata of an exemplary experiment.

@@

In a first step, two collaborative filtering techniques and two content-based filtering techniques are applied. The performance measures of the single algorithms are amended to the table. Next, aggregated user and item statistics are calculated and added to the table. Lastly, four exemplary meta-learners are applied and the performance measures of the individual approaches are added to the corpus.

@@

\subsection{Dataset Preparation}
To prepare the dataset for evaluations, duplicate interactions are merged as the recommender systems shall not recommend items to which the user has already donated to. Internal stop words in description strings of items are stripped. Furthermore, transactions containing no information about the user's location are dropped thereby enabling the consistent usage of a user's location information in new approaches.

Additionally, users having interacted less than twice are removed as well. Although, this limits the target group, the step is unavoidable as the employed validation process requires at least one interaction per user for testing and one for training. The evaluation requires user profiles and hence training-transactions in order to recommend new items which in turn need to be validated against test-transactions. Lastly, the donated amount is transformed to a transaction strength in the range of unity to five in analogy to a rating score. This step reduces the influence of outliers respectively uncommon donated amounts.

\subsection{Learning Subsystem}
The learning subsystem is a recommender system itself. SVD and KNN are used as representatives for collaborative filtering techniques and TF-IDF and word-embedding as content-based filters~\cite{scikit-learn,rehurek_lrec}. Hereby, the otherwise intrinsic inductive bias is reduced. The word-embedding relies on fastText and utilizes pre-trained vectors~\cite{DBLP:journals/corr/BojanowskiGJM16}. Vector representations of items are the normalized sum of the vectors of the normalized embeddings of the individual words. In order to work with the data in an efficient way, transactions are sampled. For this study a sample-size of $100,000$ was chosen. This represents a fair balance between required computing time and ability to generalize.

@@

A classical 5-fold cross-validation on the users is applied for training the collaborative filters. An ordering is assumed to be implied by the magnitudes of the values in the reconstructed matrix of the decomposition respectively the distances to the neighbors. The validation process of the content-based approaches uses leave-one-out. The ranking is performed using the cosine similarity of an item's vector and the user's profile.

@@

\subsection{Statistics and Application of Learning Subsystem}
The meta-learning system processes the metadata of the learning subsystem and further statistics. Its performance is evaluated via a holdout split on the transaction table using $80\%$ for training and the rest for testing. Transactions are expanded to contain more detailed information about the time, the user's location and aggregated statistics about the mean of the categories, location and interaction-strength of items with which the user has interacted with. Statistics are solely based on the training set. Values for unknown users are filled with the mean in the training data.

Four different meta-learning approaches are discussed which address the Algorithm Selection Problem (\textbf{ASP}) via a switching hybrid ensemble. The aim of the system is to predict the algorithm with the lowest recall-position for each transaction respectively each row in~\autoref{tab:IllustrativeTable}. The overall performance of a meta-learner is measured using the mean recall-position for when the recommended algorithm for each transaction is used.

@@

\begin{figure}[ht]
	\centering
	\subcaptionbox{Interaction Frequency\label{fig:InteractionFrequency}}{
		\includegraphics[width=0.45\textwidth,height=\textheight,keepaspectratio]{{{../res/number_interactions}}}
	}
	\hspace{2em}
	\subcaptionbox{Interaction Strength\label{fig:InteractionStrength}}{
		\includegraphics[width=0.45\textwidth,height=\textheight,keepaspectratio]{{{../res/distribtion_ratings}}}
	}
	\caption{The frequency of the number of interactions per user without outliers is shown in~\subref{fig:InteractionFrequency} and the distribution of the transformed interaction strength is shown in~\subref{fig:InteractionStrength}.}\label{fig:Interactions}
\end{figure}

@@

The fact that the stacking ensemble performs best is unsurprising as it is provided significantly more information than the other algorithms. The performance of the gradient booster is encouraging as it successfully addresses the ASP. From the poor performance of the user-clustering it can be reasoned that similar transactions do not necessarily yield similar results using the same recommender algorithm. Hence, the similarity of meta-feature does not imply that the performance of recommendations is similar.

@@

\section{Conclusion}
The study provides an extensively augmented dataset based on the transaction data published by DonorsChoose.org. The corpus is amended with metadata of individual recommender algorithms. Collaborative and content-based filtering techniques are used in the process and their performance is evaluated via the recall in a Top-N test set. Aggregated user and item statistics are amended to the table. Furthermore, Metadata of four switching hybrid ensemble meta-learners is amended to the dataset. The augmented public domain dataset lays the groundwork that future evaluations of existing and novel meta-learning approaches can build upon.
